---
title: "D590_Final_Project_Part_2"
author: "Brian Kenney, Prabakar Mohan, Brendan Kelly"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---


Load Library Packages

```{r setup, include=FALSE}
library(fpp3)
library(fable.prophet)
```

# 1 - Analysis of AAPL
# 1.0 - Analysis Setup

## Set up File name

```{r}
file = "Data/AAPL_2023-03-26.csv"
```

## Read the .CSV and select Date and Closing Price

```{r}
Stock_information <- read.csv(file)

selected_stock_information <- Stock_information %>% 
  select(c('Date','Close')) %>% 
  mutate(row_value = row_number())
```

## Collect Stock/ETF Ticker to be displayed on future visualizations to remove confusion

```{r}
file_name = substring(file,6,9)
```

## First set

```{r}
selected_stock_information_tsibble <- as_tsibble(selected_stock_information, index = row_value)
```

## Set up for Monthly; second set. This could be useful in the future.

```{r}
selected_stock_information_monthly <- yearmonth(selected_stock_information$`Date`, format = "%Y-%m-%d") 
selected_stock_information$`Date` <- selected_stock_information_monthly
selected_stock_information_montly_tsibble <- as_tsibble(selected_stock_information, index = row_value) 
```

```{r}
selected_stock_information_montly_aggregated_tsibble <- selected_stock_information_montly_tsibble %>% 
  aggregate(Close ~ Date, sum) %>% 
  mutate(row_value = row_number()) %>% 
  as_tsibble(index = row_value) 
```

\newpage
# 1.1 - Decomposition

```{r}
dcmp <- selected_stock_information_tsibble |>
  model(stl = STL(Close))

components(dcmp) |> autoplot() +
  labs(title = paste(file_name, "STL decomoposition"),
       y = "Stock Value ($)",
       x = "Trading Days")
```

\newpage
# 1.2 - Time Series Visualization (needs another visualization)

```{r}
autoplot(selected_stock_information_tsibble) +
  labs(y = "Stock Value ($)",
       title = paste(file_name,"Daily Closing Price over Time"),
       x = "Trading Days")
```


## Checking if the data is stationary. If kpss_stat > kpss_pvalue, then we reject null hypothesis and claim the data is non-stationary. Differencing will be required.

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_kpss)
```

## Output is the number of differences make the data stationary

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_ndiffs)
```

## The difference 

```{r}
selected_stock_information_tsibble |>
  mutate(diff_close = difference(Close)) |>
  features(diff_close, unitroot_kpss)
```


### Using Box-Cox Transformation does not change the shape of the Closing Price with a lambda value of 1.

```{r}
lambda <- selected_stock_information_tsibble |>
  features(Close, features = guerrero) |>
  pull(lambda_guerrero)
selected_stock_information_tsibble |>
  autoplot(box_cox(Close, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste(file_name,
         "Transformed Closing Price with $\\lambda$ = ",
         round(lambda,2))))
```


\newpage
# 1.3 - Description of the Time Series:

### INPUT SOMETHING 

\newpage
# 1.4 - TS Models (transformations)

### Input SOMETHING like differencing?

\newpage
# 1.5 - Predictions

## ARIMA approach

```{r}
selected_stock_information_tsibble |>
  gg_tsdisplay(difference(Close), plot_type='partial')
```

## Generate a few ARIMA orders and then compare results.

```{r}
selected_stock_fit <- selected_stock_information_tsibble |>
  model(arima010 = ARIMA(Close ~ pdq(0,1,0)),
        arima110 = ARIMA(Close ~ pdq(1,1,0)),
        arima011 = ARIMA(Close ~ pdq(0,1,1)),
        arima111 = ARIMA(Close ~ pdq(1,1,1)),
        stepwise = ARIMA(Close),
        search = ARIMA(Close, stepwise=FALSE))

print(selected_stock_fit$search)

glance(selected_stock_fit) |> arrange(AICc) |> select(.model:BIC)
```

## Show the residuals using the 'search' ARIMA 

```{r}
selected_stock_fit |>
  select(search) |>
  gg_tsresiduals()
```

### Portmanteau test shows a large P-value, suggesting the residuals are simliar to white noise

```{r}
augment(selected_stock_fit) |>
  filter(.model=='search') |>
  features(.innov, ljung_box, lag = 10, dof = 3)
```

## First Prediction using ARIMA

```{r}
selected_stock_fit |>
  forecast(h=10) |>
  filter(.model=='search') |>
  autoplot(selected_stock_information_tsibble)  +
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```

## Second approach, first creating a Train set and a Test set

```{r}
length_df = nrow(selected_stock_information_tsibble)
Train_number = round(length_df * 0.98, 0)


Train <- selected_stock_information_tsibble[0:Train_number,]
Test <- selected_stock_information_tsibble[Train_number:length_df,]

```

## Period is set to twelve for the months, and order = 1 since data is non-seasonal

```{r}
selected_stock_information_tsibble_fit <- Train |>
  model(
    arima = ARIMA(Close),
    ets = ETS(Close),
    prophet = prophet(Close ~ season(period = 12, order = 1,
                                    type = "additive"))
  )
```

## Prophet model is not good since there is no seasonality

```{r}
selected_stock_information_tsibble_fit_fc <- selected_stock_information_tsibble_fit |> forecast(h = 30)
selected_stock_information_tsibble_fit_fc |> autoplot() + autolayer(selected_stock_information_tsibble) +   
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```



```{r}

```



\newpage
# 2 - Analysis of MSFT
# 2.0 - Analysis Setup

## Set up File name

```{r}
file = "Data/MSFT_2023-03-26.csv"
```

## Read the .CSV and select Date and Closing Price

```{r}
Stock_information <- read.csv(file)
selected_stock_information <- Stock_information %>% select(c('Date','Close')) %>% 
  mutate(row_value = row_number())
```

## Collect Stock/ETF Ticker to be displayed on future visualizations to remove confusion

```{r}
file_name = substring(file,6,9)
```

## First set

```{r}
selected_stock_information_tsibble <- as_tsibble(selected_stock_information, index = row_value)
```

## Set up for Monthly; second set. This could be useful in the future.

```{r}
selected_stock_information_monthly <- yearmonth(selected_stock_information$`Date`, format = "%Y-%m-%d") 
selected_stock_information$`Date` <- selected_stock_information_monthly
selected_stock_information_montly_tsibble <- as_tsibble(selected_stock_information, index = row_value) 
```

```{r}
selected_stock_information_montly_aggregated_tsibble <- selected_stock_information_montly_tsibble %>% 
  aggregate(Close ~ Date, sum) %>% mutate(row_value = row_number()) %>% 
  as_tsibble(index = row_value) 
```

\newpage
# 2.1 - Decomposition

```{r}
dcmp <- selected_stock_information_tsibble |>
  model(stl = STL(Close))

components(dcmp) |> autoplot() +
  labs(title = paste(file_name, "STL decomoposition"),
       y = "Stock Value ($)",
       x = "Trading Days")
```

\newpage
# 2.2 - Time Series Visualization (needs another visualization)

```{r}
autoplot(selected_stock_information_tsibble) +
  labs(y = "Stock Value ($)",
       title = paste(file_name,"Daily Closing Price over Time"),
       x = "Trading Days")
```


## Checking if the data is stationary. If kpss_stat > kpss_pvalue, then we reject null hypothesis and claim the data is non-stationary. Differencing will be required.

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_kpss)
```

## Output is the number of differences make the data stationary

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_ndiffs)
```

## The difference 

```{r}
selected_stock_information_tsibble |>
  mutate(diff_close = difference(Close)) |>
  features(diff_close, unitroot_kpss)
```


### Using Box-Cox Transformation does not change the shape of the Closing Price with a lambda value of 1.

```{r}
lambda <- selected_stock_information_tsibble |>
  features(Close, features = guerrero) |>
  pull(lambda_guerrero)
selected_stock_information_tsibble |>
  autoplot(box_cox(Close, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste(file_name,
         "Transformed Closing Price with $\\lambda$ = ",
         round(lambda,2))))
```


\newpage
# 2.3 - Description of the Time Series:

### INPUT SOMETHING 

\newpage
# 2.4 - TS Models (transformations)

### Input SOMETHING like differencing?

\newpage
# 2.5 - Predictions

## ARIMA approach

```{r}
selected_stock_information_tsibble |>
  gg_tsdisplay(difference(Close), plot_type='partial')
```

## Generate a few ARIMA orders and then compare results.

```{r}
selected_stock_fit <- selected_stock_information_tsibble |>
  model(arima010 = ARIMA(Close ~ pdq(0,1,0)),
        arima110 = ARIMA(Close ~ pdq(1,1,0)),
        arima011 = ARIMA(Close ~ pdq(0,1,1)),
        arima111 = ARIMA(Close ~ pdq(1,1,1)),
        stepwise = ARIMA(Close),
        search = ARIMA(Close, stepwise=FALSE))

print(selected_stock_fit$search)

glance(selected_stock_fit) |> arrange(AICc) |> select(.model:BIC)
```

## Show the residuals using the 'search' ARIMA 

```{r}
selected_stock_fit |>
  select(search) |>
  gg_tsresiduals()
```

### Portmanteau test shows a large P-value, suggesting the residuals are simliar to white noise

```{r}
augment(selected_stock_fit) |>
  filter(.model=='search') |>
  features(.innov, ljung_box, lag = 10, dof = 3)
```

## First Prediction using ARIMA

```{r}
selected_stock_fit |>
  forecast(h=10) |>
  filter(.model=='search') |>
  autoplot(selected_stock_information_tsibble)  +
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```

## Second approach, first creating a Train set and a Test set

```{r}
length_df = nrow(selected_stock_information_tsibble)
Train_number = round(length_df * 0.98, 0)


Train <- selected_stock_information_tsibble[0:Train_number,]
Test <- selected_stock_information_tsibble[Train_number:length_df,]

```

## Period is set to twelve for the months, and order = 1 since data is non-seasonal

```{r}
selected_stock_information_tsibble_fit <- Train |>
  model(
    arima = ARIMA(Close),
    ets = ETS(Close),
    prophet = prophet(Close ~ season(period = 12, order = 1,
                                    type = "additive"))
  )
```

## Prophet model is not good since there is no seasonality

```{r}
selected_stock_information_tsibble_fit_fc <- selected_stock_information_tsibble_fit |> forecast(h = 30)
selected_stock_information_tsibble_fit_fc |> autoplot() + autolayer(selected_stock_information_tsibble) +   
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```




```{r}

```

\newpage
# 3 - Analysis of TSLA
# 3.0 - Analysis Setup

## Set up File name

```{r}
file = "Data/TSLA_2023-03-26.csv"
```

## Read the .CSV and select Date and Closing Price

```{r}
Stock_information <- read.csv(file)

selected_stock_information <- Stock_information %>% 
  select(c('Date','Close')) %>% 
  mutate(row_value = row_number())
```

## Collect Stock/ETF Ticker to be displayed on future visualizations to remove confusion

```{r}
file_name = substring(file,6,9)
```

## First set

```{r}
selected_stock_information_tsibble <- as_tsibble(selected_stock_information, index = row_value)
```

## Set up for Monthly; second set. This could be useful in the future.

```{r}
selected_stock_information_monthly <- yearmonth(selected_stock_information$`Date`, format = "%Y-%m-%d") 
selected_stock_information$`Date` <- selected_stock_information_monthly
selected_stock_information_montly_tsibble <- as_tsibble(selected_stock_information, index = row_value) 
```

```{r}
selected_stock_information_montly_aggregated_tsibble <- selected_stock_information_montly_tsibble %>% 
  aggregate(Close ~ Date, sum) %>% 
  mutate(row_value = row_number()) %>% 
  as_tsibble(index = row_value) 
```

\newpage
# 3.1 - Decomposition

```{r}
dcmp <- selected_stock_information_tsibble |>
  model(stl = STL(Close))

components(dcmp) |> autoplot() +
  labs(title = paste(file_name, "STL decomoposition"),
       y = "Stock Value ($)",
       x = "Trading Days")
```

\newpage
# 3.2 - Time Series Visualization (needs another visualization)

```{r}
autoplot(selected_stock_information_tsibble) +
  labs(y = "Stock Value ($)",
       title = paste(file_name,"Daily Closing Price over Time"),
       x = "Trading Days")
```


## Checking if the data is stationary. If kpss_stat > kpss_pvalue, then we reject null hypothesis and claim the data is non-stationary. Differencing will be required.

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_kpss)
```

## Output is the number of differences make the data stationary

```{r}
selected_stock_information_tsibble |>
  features(Close, unitroot_ndiffs)
```

## The difference 

```{r}
selected_stock_information_tsibble |>
  mutate(diff_close = difference(Close)) |>
  features(diff_close, unitroot_kpss)
```


### Using Box-Cox Transformation does not change the shape of the Closing Price with a lambda value of 1.

```{r}
lambda <- selected_stock_information_tsibble |>
  features(Close, features = guerrero) |>
  pull(lambda_guerrero)
selected_stock_information_tsibble |>
  autoplot(box_cox(Close, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste(file_name,
         "Transformed Closing Price with $\\lambda$ = ",
         round(lambda,2))))
```


\newpage
# 3.3 - Description of the Time Series:

### INPUT SOMETHING 

\newpage
# 3.4 - TS Models (transformations)

### Input SOMETHING like differencing?

\newpage
# 3.5 - Predictions

## ARIMA approach

```{r}
selected_stock_information_tsibble |>
  gg_tsdisplay(difference(Close), plot_type='partial')
```

## Generate a few ARIMA orders and then compare results.

```{r}
selected_stock_fit <- selected_stock_information_tsibble |>
  model(arima010 = ARIMA(Close ~ pdq(0,1,0)),
        arima110 = ARIMA(Close ~ pdq(1,1,0)),
        arima011 = ARIMA(Close ~ pdq(0,1,1)),
        arima111 = ARIMA(Close ~ pdq(1,1,1)),
        stepwise = ARIMA(Close),
        search = ARIMA(Close, stepwise=FALSE))

print(selected_stock_fit$search)

glance(selected_stock_fit) |> arrange(AICc) |> select(.model:BIC)
```

## Show the residuals using the 'search' ARIMA 

```{r}
selected_stock_fit |>
  select(search) |>
  gg_tsresiduals()
```

### Portmanteau test shows a large P-value, suggesting the residuals are simliar to white noise

```{r}
augment(selected_stock_fit) |>
  filter(.model=='search') |>
  features(.innov, ljung_box, lag = 10, dof = 3)
```

## First Prediction using ARIMA

```{r}
selected_stock_fit |>
  forecast(h=10) |>
  filter(.model=='search') |>
  autoplot(selected_stock_information_tsibble)  +
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```

## Second approach, first creating a Train set and a Test set

```{r}
length_df = nrow(selected_stock_information_tsibble)
Train_number = round(length_df * 0.98, 0)


Train <- selected_stock_information_tsibble[0:Train_number,]
Test <- selected_stock_information_tsibble[Train_number:length_df,]

```

## Period is set to twelve for the months, and order = 1 since data is non-seasonal

```{r}
selected_stock_information_tsibble_fit <- Train |>
  model(
    arima = ARIMA(Close),
    ets = ETS(Close),
    prophet = prophet(Close ~ season(period = 12, order = 1,
                                    type = "additive"))
  )
```

## Prophet model is not good since there is no seasonality

```{r}
selected_stock_information_tsibble_fit_fc <- selected_stock_information_tsibble_fit |> forecast(h = 30)
selected_stock_information_tsibble_fit_fc |> autoplot() + autolayer(selected_stock_information_tsibble) +   
  labs(y = "Stock Value ($)",
       title = paste(file_name, "Daily Closing Price over Time"),
       x = "Trading Days")
```



```{r}

```


# Team contributions (pending...)

Brian: Completed rough draft of the Part 2 submission. 

Mohan: Generated Github repo and place Datasets. Picked option 1.

Brendan: Picked option 1.